---
title: "Homework 10"
author: "Caleb Fant"
date: "2024-04-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

UT EID: cwf678

Github Link: https://github.com/CaleFan05/SDS315.git

```{r}
library(tidyverse)
library(moderndive)
library(effectsize)
```

## Problem 1: Redlining

### Question:

Was there an association between the number of FAIR policies and the racial/ethnic composition of a zip code in Chicago in the 1970s when adjusting for rates of fire, homes built before World War II, and median family income?

### Approach:

We performed a linear least squares regression on a dataset with relevant information on Chicago zip codes in the 1970s using policies per 100 housing units as the response variable, and minority percentage, houses built before WWII percentage, median family income, and fires per 100 houses as the predictor variables. This was done with the "lm" function in R.

### Results:

```{r}
redline = read_csv("redlining.csv")
# lm(policies ~ minority, data=redline) %>% get_regression_table(conf.level = 0.95, digits=3)
lm(policies ~ minority + fire + age + income, data=redline) %>% get_regression_table(conf.level = 0.95, digits=3)
```

### Conclusion:

Even after adjusting for fire, age, and income, we can still say with 95% confidence that an additional 1% of minorities in a zip code is correlated with between .003 and .014 additional FAIR plan policies/renewals per 100 housing units.

## Problem 2:
```{r}
groceries = read_csv("groceries.csv")
```

### Part A.
```{r}
price_means = groceries %>%
  group_by(Store) %>%
  summarise("mean_price"=mean(Price))
ggplot(price_means, aes(Store,mean_price)) + geom_col() + coord_flip() + labs(title="Bar Chart of Prices at Grocery Stores", y="Average Price of Products Sold")
```
<br />This plot shows the how the average price of products differs from store to store in the dataset.

### Part B.

```{r}
# product_counts = groceries %>%
#   group_by(Product) %>%
#   summarise("Count" = n())
# ggplot(product_counts,aes(Product,Count)) + geom_col() + coord_flip()
ggplot(groceries,aes(x=Product)) + geom_bar() + coord_flip() + labs(x="Product",y="Number of Stores",title="Bar Chart of the Number of Stores Selling Each Product")
```
<br />This plot shows the differences in availability between products in the dataset.

### Part C.
```{r}
c_table = lm(Price ~ Product + Type, data=groceries) %>% get_regression_table(conf.level = .95,digits=2)
c_table %>% rmarkdown::paged_table()
```
<br />Compared with ordinary grocery stores (like Albertsons, HEB, or Krogers), convenience stores charge somewhere between 0.41 and 0.92 dollars more for the same product.

### Part D.
```{r}
lm(Price ~ Product + Store, data=groceries) %>% get_regression_table(conf.level = .95,digits=2) %>% rmarkdown::paged_table()
```
<br />Walmart and Kroger Fresh Fare appear to charge the least for the same product, while Wheatsville and Whole Foods appear to charge the most.

### Part E.

I would argue that Central Market charges a similar amount to HEB for the same product. We are 95% confident that purchasing a product at Central Market will cost between 23 to 92 cents less than what you would pay at Albertson's. For HEB, this interval is between 35 and 95 cents. Considering the large overlap between these two intervals, I would argue that there is not evidence supporting the theory that Central Market charges more than HEB for the same product.

### Part F.
```{r}
groceries = groceries %>% mutate(Income10k = Income/10000)
lm(Price ~ Product + Income10k, data = groceries) %>% get_regression_table(conf.level = 0.95, digits = 3) %>% rmarkdown::paged_table()
lm(Price ~ Product + Income10k, data = groceries) %>% standardize_parameters()
```
<br />Consumers in poorer zip codes seem to pay more for the same product on average. The first table shows how much we should expect the price to change for a one unit increase in the predictor variable. In this case, it shows a -0.014 change in price for a 10k dollar increase in average annual income for a zip code. Because this number is negative it suggests a decrease in price is associated with a higher income. We can also use the second table to see that a one-standard deviation increase in the income of a ZIP code seems to be associated with a -0.03 standard-deviation change in the price that consumers in that ZIP code expect to pay for the same product.